Can we use masked priming in an online setting?

### 2021-05-19 MaskedPriming_complete_mspg.Rmd
Implemented Pablo's changes

  We thank Grace Backhouse for her help with data collection. The data for Experiment 1 were collected as part of her MSc dissertation.
  This research has been partly supported by Grant PSI2017-86210-P  from  the  Spanish Ministry of Science and Innovation (MP), Grant PGC2018-097145-B-I00 from the Spanish State Research Agency (AB), \color{red} and... PABLO?  \color{black} 
abstract: |
  Masked priming is one of the most important paradigms in the study of visual word recognition, but it is usually thought to require a laboratory setup with a known monitor and keyboard. To test if this technique can be safely used in an online setting, we conducted two online masked priming lexical decision task experiments using PsychoPy/PsychoJS [@peirce2019]. Importantly, we also tested the role of prime exposure duration (33.3 vs. 50 ms in Experiment 1 and 16.7 vs. 33.3 ms in Experiment 2), thus allowing us to examine both across conditions and within-conditions effects. We found that our online data are indeed very similar to the masked priming data reported by @gomezDiffusionModelAccount2013. Additionally, we found a clear effect of prime duration, with the priming effect (measured in terms of response time and accuracy) being stronger at 50 ms than 33.3 ms and no priming effect at 16.7 ms prime duration. From these results, we can conclude that modern online browser-based experimental psychophysics packages (e.g., PsychoPy) can present stimuli and collect responses on standard end user devices with enough precision. These findings provide us with confidence that masked priming can be used online, thus allowing us not only to run less time-consuming experiments, but also to reach populations that are difficult to test in a laboratory.

@@

Cognitive psychologists have been much slower that their behavioral economics coleagues in taking up online paradigms [see @brysbaert2016impact; @cai2017accent;@eerland2013influence;@Dufau_2011; @rodd2016impact, for some exceptions], often due to concerns about the validity of the results. Such concerns are not limited to cognitive studies [see @austSeriousnessChecksAre2013], but they are exacerbated by the reliance on precise presentation times in cognitive psychology. Furthermore, these concerns are even more central in the masked priming technique, where it is critical for the onset of the mask, the prime, and the target to occur at the nominal times. For instance, presenting the masked prime for longer than intended (e.g., 83 ms or longer instead of the nominal 50 ms) could counteract the effect of the mask, making the prime consciously visible to the participant and possibly altering the processes of interest [see @Zimmerman_2012].

@@

Specifically, in this study we were interested in whether we could replicate and extend a key phenomenon in laboratory masked priming lexical decision using an online setup: masked identity priming is commonly described as a savings effect. As first suggested by @forster1998pros, for a masked identity prime, "the lexical entry is already in the process of being opened, and hence the evaluation of this entry begins sooner", whereas for an unrelated prime, "the entry for the target word would be closed down (since it fails to match the prime), and no savings would occur" (p. 213). Thus, according to the savings account, a target word like *DOCTOR* would enjoy an encoding advantage when preceded by an identity prime such as *doctor* than when preceded by an unrelated prime such as *pencil* (i.e., a head-start). One implication of such benefit is that the RT distributions of the unrelated and identity pairs should reflect a shift rather than a change in shape. Furthermore, this shift should be approximately similar in magnitude to the prime-target stimulus-onset asynchrony (SOA). Empirical evidence supporting this view has been obtained in several papers not only with skilled adults but also with developing readers [e.g., @gomezDiffusionModelAccount2013; @Gomez_2020; @Taikh_2020; @Yang_2021]. Moreover, this pattern of results fits very well with the diffusion model [@ratcliff2004]. This model proposes that, when making a two-choice decision, the resultant RT can be explained as the sum of non-decision parameters, which are the encoding time and response execution ($T_{er}$) and decision parameters, which refer to the process of accumulation of information until a decision criteria is reached. Importantly, in the decision process, the information gathered from the stimulus can vary in noise, depending on its quality, which modifies the rate at which information is accumulated (i.e., the *drift rate*). With regards to RTs from masked priming tasks, @gomezDiffusionModelAccount2013 found that the difference between identity and unrelated conditions could be accounted by a change in the $T_{er}$ parameter, while there were no differences across conditions in the parameter that corresponds to the quality of evidence gathered (i.e., drift rate)--note that changes in drift rate would necessarily produce a skewer RT distribution in the slower, unrelated condition.

\color{green} I don't love the prior paragrpah.  My main concenrn is that the savings account could have been also implemented as a headstart that boosts the drift rate.  The Ter model is just one implementation of the hard start hypothesis  \color{black}

@@

In Experiment 2, we found the pattern we expected from previous research: In the mean $\mu$ of the exGaussian distribution, we observed a robust identity priming effect in response time and accuracy in the 33.3 ms prime duration, but a very weak effect in the 16.7 ms prime duration [@Tzur_2007; @ziegler2000]. This outcome suggests that the timing in our online experiments was likely to be quite close to the timing set in the experiment script. In addition, we did not observe an effect of the prime relatedness on the shape parameter $\beta$ of the exponential distribution [i.e., the effect corresponded to a shift of the RT distributions, as shown in in-lab studies; e.g., @Taikh_2020; @Yang_2021]. Moreover, the magnitude of the priming effect at the 33.3 ms prime exposure duration was close to the magnitude of the prime-target stimulus-onset asynchrony, as can be seen in the delta plots in Figure \ref{fig:deltas-e2-words} [see @gomezDiffusionModelAccount2013; @Perea_2018].  Hence, we replicated again the commonly observed savings effect [@forster1998pros].

\color{green}I changed gaussian to exgaussian above, is that correct? \color{black}

>>>>

  We thank Grace Backhouse for her help with data collection. The data for Experiment 1 were collected as part of her MSc dissertation.
  This research has been partly supported by Grant PSI2017-86210-P from  the  Spanish Ministry of Science and Innovation (MP), Grant PGC2018-097145-B-I00 from the Spanish State Research Agency (AB), and internal funding from Bournemouth University (BA).
abstract: |
  Masked priming is one of the most important paradigms in the study of visual word recognition, but it is usually thought to require a laboratory setup with a known monitor and keyboard. To test if this technique can be safely used in an online setting, we conducted two online masked priming lexical decision task experiments using PsychoPy/PsychoJS [@peirce2019]. Importantly, we also tested the role of prime exposure duration (33.3 vs. 50 ms in Experiment 1 and 16.7 vs. 33.3 ms in Experiment 2), thus allowing us to examine both across conditions and within-conditions effects. We found that our online data are indeed very similar to the masked priming data reported by @gomezDiffusionModelAccount2013. Additionally, we found a clear effect of prime duration, with the priming effect (measured in terms of response time and accuracy) being stronger at 50 ms than 33.3 ms and no priming effect at 16.7 ms prime duration. From these results, we can conclude that modern online browser-based experimental psychophysics packages (e.g., PsychoPy) can present stimuli and collect responses on standard end user devices with enough precision. These findings provide us with confidence that masked priming can be used online, thus allowing us not only to run less time-consuming experiments, but also to reach populations that are difficult to test in a laboratory.

@@

Cognitive psychologists have been much slower that their behavioral economics coleagues in taking up online paradigms [see @brysbaert2016impact; @cai2017accent;@eerland2013influence; @rodd2016impact, for some exceptions], often due to concerns about the validity of the results. Such concerns are not limited to cognitive studies [see @austSeriousnessChecksAre2013], but they are exacerbated by the reliance on precise presentation times in cognitive psychology. Furthermore, these concerns are even more central in the masked priming technique, where it is critical for the onset of the mask, the prime, and the target to occur at the nominal times. For instance, presenting the masked prime for longer than intended (e.g., 83 ms or longer instead of the nominal 50 ms) could counteract the effect of the mask, making the prime consciously visible to the participant and possibly altering the processes of interest [see @Zimmerman_2012].

@@

Specifically, in this study we were interested in whether we could replicate and extend a key phenomenon in laboratory masked priming lexical decision using an online setup: masked identity priming is commonly described as a savings effect. As first suggested by @forster1998pros, for a masked identity prime, "the lexical entry is already in the process of being opened, and hence the evaluation of this entry begins sooner", whereas for an unrelated prime, "the entry for the target word would be closed down (since it fails to match the prime), and no savings would occur" (p. 213). Thus, according to the savings account, a target word like *DOCTOR* would enjoy an encoding advantage when preceded by an identity prime such as *doctor* than when preceded by an unrelated prime such as *pencil* (i.e., a head-start). One implication of such benefit is that the RT distributions of the unrelated and identity pairs should reflect a shift rather than a change in shape. Furthermore, this shift should be approximately similar in magnitude to the prime-target stimulus-onset asynchrony (SOA). Empirical evidence supporting this view has been obtained in several papers not only with skilled adults but also with developing readers [e.g., @gomezDiffusionModelAccount2013; @Gomez_2020; @Taikh_2020; @Yang_2021]. @gomezDiffusionModelAccount2013 proposed an implementation of this hypothesis within the diffusion model [@ratcliff2004]. This implementation proposes that, when making a two-choice decision, the resultant RT can be explained as the sum of non-decision parameters, which are the encoding time and response execution ($T_{er}$) and decision parameters, which refer to the process of accumulation of information until a decision criteria is reached. Importantly, in the decision process, the information gathered from the stimulus can vary in noise, depending on its quality, which modifies the rate at which information is accumulated (i.e., the *drift rate*). With regards to RTs from masked priming tasks, @gomezDiffusionModelAccount2013 found that the difference between identity and unrelated conditions could be accounted for by a change in the $T_{er}$ parameter, while there were no differences across conditions in the parameter that corresponds to the quality of evidence gathered (i.e., drift rate)--note that changes in drift rate would necessarily produce a more skewed RT distribution in the slower, unrelated condition.

@@

In Experiment 2, we found the pattern we expected from previous research: In the mean $\mu$ of the Gaussian component of the ex-Gaussian distribution, we observed a robust identity priming effect in response time and accuracy in the 33.3 ms prime duration, but a very weak effect in the 16.7 ms prime duration [@Tzur_2007; @ziegler2000]. This outcome suggests that the timing in our online experiments was likely to be quite close to the timing set in the experiment script. In addition, we did not observe an effect of the prime relatedness on the shape parameter $\beta$ of the exponential component [i.e., the effect corresponded to a shift of the RT distributions, as shown in in-lab studies; e.g., @Taikh_2020; @Yang_2021]. Moreover, the magnitude of the priming effect at the 33.3 ms prime exposure duration was close to the magnitude of the prime-target stimulus-onset asynchrony, as can be seen in the delta plots in Figure \ref{fig:deltas-e2-words} [see @gomezDiffusionModelAccount2013; @Perea_2018].  Hence, we replicated again the commonly observed savings effect [@forster1998pros].
